### 2025-05-14 manuscript/fig_captions.Rmd
change manuscript format to Fisheries Research

---
title             : "Application of Close-kin mark–recapture (CKMR) to Protogynous Fishes"
shorttitle        : "\\mbox{}"

author: 
  - name          : "Claudia Friess"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : ""
    email         : "elasmophile@gmail.com"
    role:         
  - name          : "Eric C. Anderson"
    affiliation   : "2,3"
    role:
  - name          : "Lisa E. Ailloud"
    affiliation   : "4"
    role:
  - name          : "David S. Portnoy"
    affiliation   : "5"
    role:
  - name          : "Susan K. Lowerre-Barbieri"
    affiliation   : "1,6"
    role:

affiliation:
  - id            : "1"
    institution   : "Florida Fish and Wildlife Research Institute, St. Petersburg, FL, USA"
  - id            : "2"
    institution   : "NOAA, National Marine Fisheries Service, Southwest Fisheries Science Center, Fisheries Ecology Division, Santa Cruz, CA, USA"
  - id            : "3"
    institution   : Deptartment of Biology, Colorado State University, Fort Collins, CO, USA
  - id            : "4"
    institution   : "NOAA, National Marine Fisheries Service, Southeast Fisheries Science Center, Charleston, SC, USA"
  - id            : "5"
    institution   : "Marine Genomics Laboratory, Texas A&M University, Corpus Christi, Corpus Christi, TX, USA"
  - id            : "6"
    institution   : "University of Florida, Fisheries and Aquatic Sciences, Gainesville, FL, USA"

authornote: ""

abstract: |
  Abundance estimates for marine fish are difficult to obtain and typically rely on 
  stock assessment models that assume only females matter to productivity. Many fished 
  species, however, are sequential hermaphrodites for which fishing can drastically 
  skew sex ratios, raising concerns about sustainability, especially if managed based 
  on female-only reference points. There is interest in integrating Close-Kin Mark-Recapture 
  (CKMR) into stock assessments to reduce uncertainty in abundance estimates, but this 
  has not previously been done for sequential hermaphrodites. Here we derive the CKMR kinship probabilities 
  for protogynous hermaphrodites and include them in an integrated stock assessment framework
  for estimating abundance. Using simulation, we evaluate the robustness of this approach to 
  uncertainty in male contribution to reproductive success and genetic data availability. 
  We show that CKMR can produce unbiased abundance and sex ratio estimates but only 
  if the relative contribution of male and female age classes to reproductive success 
  are not misspecified in the estimation model and the level of mitochondrial haplotype diversity 
  in the population is well estimated. We also highlight the potential for CKMR data to estimate 
  parameters of the sex transition and male reproductive contribution functions in an 
  integrated modeling framework.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "simulation, mtDNA, reproductive resilience, integrated stock assessment, hermaphrodites, genetic tagging"

bibliography      : ["citations.bib", "r-references.bib"]
csl               : "cjfas.csl"

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

@@

classoption       : "man"
output            : papaja::apa6_pdf

@@

@@

  - \captionsetup[figure]{labelfont=bf}
  - \captionsetup[table]{labelfont=bf, labelsep = period, textformat = simple}
  - \raggedbottom

@@

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
library(tidyverse)
library(stargazer)
library(kableExtra)
library(gridExtra)
library(nplyr)
library(patchwork)
library(ggh4x)
library(cowplot)
library(gridExtra)
library(grid)
library(scales)
library(here)

# caching for quick knitting.  
#    set cache = FALSE to not cache
#    set cache.extra to a new value to remove all caches and re-evaluate
knitr::opts_chunk$set(cache.extra = 4, cache = FALSE)

```

```{r figures, include = FALSE}
source(here("code", "analysis", "results_tables.R"), local = knitr::knit_global())
source(here("code", "analysis", "results_figures.R"), local = knitr::knit_global())
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

Generating accurate estimates of marine fish population abundance is one of the most challenging 
yet essential mandates of fisheries science. To meet it, ever-more sophisticated stock assessment
models, capable of integrating a variety of data sources, are being developed [@punt2020essential].
At the core, however, these models are based on the same basic fish population dynamics 
equations developed for gonochoristic species decades ago [@beverton1993dynamics; @hilborn1992quantitative]. Similarly unchanging
has been the central overarching goal of fisheries management: optimize resource exploitation by balancing current 
extraction against future productivity, or reproductive potential. 
The key, then, is how we define, both conceptually and operationally, "reproductive potential". In terrestrial mammals, it is driven by the 
number and fecundity of the females. This paradigm has been adopted for marine fish in the form of measures
like spawning stock biomass (SSB) or total egg production, although problems with this approach are increasingly 
recognized [@sharma2019recruitment]. The reproductive resilience paradigm posits that marine fish reproductive strategies 
are unlike those seen on land and evolved to achieve lifetime reproductive success and maintain population 
stability within highly variable environments [@tringali2023reproductive; @lowerre2017reproductive]. 
These reproductive strategies include extreme fecundity, external fertilization of pelagic eggs, 
offspring survivorship affected by where and when fish spawn, and sequential hermaphroditism.
Both protandry (reproducing as males first and then as females) and protogyny (reproducing as females first and changing sex to male) 
occur in marine fishes, but protogyny is most common [@pla2022switches], especially in warmwater
systems [@lowerre2023unified].

Delayed male maturation, such as seen in protogynous species, is predicted to evolve when there
is male competition for females. Sex allocation theory predicts that adult sex ratios are driven
by processes that result in optimal mean reproductive success for the population [@charnov1982theory; @warner1988sex] and the size 
advantage model predicts that a fish will change sex when reproductive success at age is greater for
the terminal sex [@ghiselin1969evolution; @warner1975reproductive; @munoz2003new]. 
Fisheries target larger animals, truncating population age and size structure [@barnett2017old], 
which disproportionately increases the risk of overexploitation for males in protogynous species.
Yet, many protogynous species support important fisheries, 
such as the keystone predator California sheephead, *Bodianus pulcher*, [@alonzo2004status; @hamilton2015exploitation] and 
heavily-fished serranid groupers in the Gulf of Mexico (GOM) and Caribbean Sea.

Two current areas of active research in fisheries stock assessment are defining and measuring 
reproductive potential for protogynous hermaphrodites [@alonzo2008importance; @brooks2008stock;
@alonzo2005sex] and incorporating genetic data into integrated models [@punt2020essential; @punt2024including].
Of particular interest is Close-Kin Mark-Recapture (CKMR) [@nielsen2001statistical;
@julius2001allele; @bravington2016close], a method that
uses genetic data to identify kin pairs in a sample and treats those
kin pairs as "recaptures" in a mark-recapture-like framework [see @casas2023review for a 
review of CKMR and its application to fisheries].
CKMR has been used to estimate absolute abundance for a number of marine and freshwater fish 
species with a range of reproductive strategies, including the
gonochoristic southern Bluefin tuna, *Thunnus thynnus*, [@bravington2016absolute; @davies2020next], 
semelparous Chinook salmon, *Oncorhynchus tshawytscha*, [@rawding2014genetic], and elasmobranchs which have a fixed-ratio 
reproductive strategy [@hillary2018genetic; @delaval2023evaluating].
Integrating CKMR into a stock assessment requires including the CKMR pseudolikelihood
[@bravington2016close] as one component in the stock assessment model.
This CKMR pseudolikelihood relates the observed number of genetically identified closely related pairs
(e.g., parent-offspring or half-sibling-pairs) in a sample to the
number expected given the size of the population. The expected
number of kin pairs is found by deriving a set of _kinship probabilities_, each one being
the probability that a pair of samples are close kin, given the size of the population and
additional information that describes the likelihood of an individual surviving and
reproducing over time relative to other individuals with similar demographic characteristics.

To date, nearly all CKMR applications have been to gonochoristic species and have 
assumed fecundity drives reproductive success and that this success scales
equally between males and females. The life history of protogynous hermaphrodites 
complicates application of CKMR because 
sex ratios are typically skewed toward females, and, amongst half-sibling
pairs, one fish's mother could be another's father. In fact, CKMR kinship probabilities 
for protogynous hermaphrodites have not previously been derived. Additionally, it is 
unknown whether traditionally-used heuristics for the CKMR sample sizes required
to produce accurate and precise abundance estimates in gonochoristic species are 
also appropriate for protogynous hermaphrodites. Finally, even though current best
practice in stock assessments of protogynous hermaphrodites is to assume male and
female SSB contribute equally to reproductive potential [@brooks2008stock], it has not been determined 
whether this is, in fact, the case, nor do we know if and how a declining male sex ratio
affects population productivity.

In this study, we derive the kinship probabilities for protogynous hermaphrodites,
extending the work of @bravington2016close to use the frequency of kinship types 
observed in genetic data to estimate abundance and demographics in protogynous 
species. We then conduct a simulation study
to investigate the robustness of these estimators to various levels of genetic information availability
as well as uncertainty in male contribution to reproductive success and key reproductive and genetic parameters. 
The simulation consists of an individual-based operating model (OM) that generates pedigrees for a population 
of protogynous hermaphrodites experiencing increasing levels of fishing mortality 
and concurrent population decline and increased sex-ratio skew. Each simulation creates millions of
fish _in silico_ with known birth and death dates, sex, and mortality causes, and a known pedigree 
that connects them all. We then sample individuals from the OM using different data collection 
strategies—varying both the number of years and the sample size—to provide both traditional fisheries data and
CKMR kin pairs for analysis in a simple, integrated fisheries stock-assessment-style estimation 
model (EM). Finally, we quantify the bias in estimated abundance and male sex ratio 
under different OM configurations regarding male reproductive success and when key genetic and reproductive 
parameters or quantities are treated as known, estimated, or misspecified in the EM. 
Of particular interest was whether there is sufficient 
information in the CKMR data to estimate the shapes of the sex transition and male age-specific reproductive
output functions used in the CKMR likelihood.

# Methods

This section is organized as follows: 
we first describe the CKMR kinship probabilities for protogynous hermaphrodites, then
present the different components of the simulation study to test this model: the (individual-based) 
operating model (OM) to simulate 'reality', the sampling model to generate 'observed' data from 
the OM, and finally the integrated fisheries stock assessment estimation model (EM) that uses
the 'observed' data and the derived CKMR kinship probabilities to estimate a set of parameters 
that allow generating 'estimated' abundance which can then be compared to the 'true' abundance 
generated by the OM. For each simulation component, we describe the configurations or scenarios 
we used to test the performance of our CKMR estimators (Figure \@ref(fig:simoverview-plot)).

## CKMR Model

There are several close kin relationships that could be included. For simplicity, we
limit this study to half-sibling pairs (HSPs) and parent-offspring pairs (POPs) and
assume that these can be identified without error. We furthermore assume that ages of 
sampled individuals are known without error, sampling is conducted on dead
individuals, and age-specific survival is independent of sex. We assume the species is monandric (i.e., all males
originate from females) and that sex change, if it occurs, happens at the
beginning of the annual interval, followed by reproduction, then mortality/fishing/sampling. 
The full CKMR model derivation can be found in Appendix A. The probabilities of observing
HSPs (Equations A1-A7) account for the unknown age and sex of the shared parent as
well as the possibility that the first-born's mother could be the second-born's father.
Similarly, the probabilities of observing POPs (Equations A9-A13) account for the fact that
the parent, if it was male at the time of sampling, could have been female when it
gave birth to the offspring in the pair. Sex change is included via the age-specific
probability of a female changing sex to become male, $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
(see Table \@ref(tab:table1) for this and other notation used in this paper).

The ability to identify---or at least infer---the sex of a sampled individual's parent is crucial for 
applying the model, and in our model this is accomplished by sequencing the fish's
mitochondrial DNA (mtDNA) which is inherited from the female parent. For the purposes
of the simulation analyses, we assumed that different mtDNA haplotypes were equifrequent,
so that we were able to parameterize the mtDNA haplotype frequencies simply by $H_\mathrm{mt}$---the
probability that two randomly sampled mtDNAs from the population
are different in sequence. This quantity is also commonly called the mtDNA
haplotype diversity [@avise2012molecular], and parameterizing the simulations in terms of it allowed
us to avoid simulating mtDNA sequences
in all simulated individuals and also simplified the CKMR likelihood used to analyze the simulations as described
in the Supplement. An $\Hmt$ of 1 would indicate a 100% certainty
that two HSPs with the same mtDNA sequence shared a mother, while any value <1 means there is
a chance (which equals $1 - \Hmt$) that their mtDNA sequence is the same even though they did not
share a mother. We incorporate uncertainty due to $\Hmt$ into our CKMR model
equations (Equations A8, and A14-A17) and in the simulation, we explore the effects of 
misspecifying both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\Hmt$ in the estimation model (see below).

## Model Species

We parameterized our simulation model based on life history data for Gag, 
*Mycteroperca microlepis*, in the GOM. Gag live to be over 30 years old and 
mature around age 4. Gag exhibit clear depth preferences with life stage, sex, 
and spawning, utilizing estuarine nursery habitat as young of the year, 
moving to nearshore sites with structure as young females (~ages 1 to 4), 
and then moving offshore to spawn as they mature. Individuals undergoing sex transition occur before, 
during, and after the spawning season and in both nearshore all-female 
pre-spawning aggregations and in offshore spawning grounds [@lowerre2020testing].
Gag spawn at the shelf edge from February to mid-April, likely in pairs.
They produce pelagic eggs that settle in estuaries after 35 to 45 days [@fitzhugh2005spatial].
Males are thought to remain in deeper waters on the shelf edge year-round whereas some females return to 
nearshore structure after the spawning season. Although a male as young as four years 
old has been observed, age at 50% male is approximately 12 years old.

Gag are highly sought after in both the commercial and recreational fisheries in 
the GOM. The majority of the catch quota is allocated to the recreational 
sector, and estimates of recreational landings are considered highly uncertain [@SEDAR2022; @NOAA2023]. 
Gag have been estimated to be overfished and experiencing overfishing almost 
consistently since the late 1960’s [@SEDAR2022]. The spawning stock biomass estimated 
in the 2022 stock assessment was about 2% of unfished levels, and the sex ratio has declined from an 
estimated 32% male in unfished conditions to between 1 and 5% male in recent years 
[@lowerre2020testing; @SEDAR2022]. It is unknown to what extent sperm limitation 
is responsible for the continued decline of the gag population [@lowerre2021gag].

## Operating Model

The OM is a discrete-time, individual-based model (IBM) implemented in R based on the 'fishSim' 
package [@R-fishSim]. We modified the package functions to accommodate 
our study needs (e.g., by adding the ability to model protogynous hermaphrodite 
life histories) and rewrote the slower-running R functions in C++ using the 'Rcpp' 
package [@eddelbuettel2013seamless] to increase computational efficiency. The basic
annual sequence, modeled loosely on gag life history,
was that sex transition and then spawning occurred at the beginning
of the year, followed by mortality, CKMR sampling, and age incrementation (Figure \@ref(fig:flowchart-plot)).
For computational efficiency, the simulation starts at age 1, given that most age-0 gag
are not available to the fishery and suffer high rates of natural mortality.

Simulated reproductive value for females and males of age $a$, denoted as  $\mathrm{RV}_{a,F}$ and
$\mathrm{RV}_{a,M}$, respectively, were used to produce age-1 offspring and assign parentage.
$\mathrm{RV}_{a,F}$ represents the age-specific expected number 
of age-1 offspring produced per female. This was modeled as a power function 
(Figure \@ref(fig:RC-plot), panel A) derived 
from female gag batch fecundity-at-age data, and scaled to achieve a stable population in 
the absence of fishing. To help produce recruitment variability, $\mathrm{RV}_{a,F}$ 
was randomly scaled up or down annually by sampling a reproductive value, 
$\mathrm{RV}_{a,t,F,b}$ for each breeding group, $b$, from a $\mathrm{Uniform}(0,1)$ 
distribution and rescaling such that $\sum_a{\mathrm{RV}_{a,F}e^{\eta_t}} = \sum_a{\sum_b\mathrm{RV}_{a,t,F,b}}$.
Here, $\eta_t$ represents the annual scaling factor, ensuring that the total RV 
across all ages and breeding groups remains consistent.
The realized number of age-1 offspring in a given year $t$
for a given female of age $a$ in breeding group $b$ was a Poisson random draw from 
$\mathrm{RV}_{a,t,F,b}$. All males were considered mature, and $\mathrm{RV}_M$ 
was used as the age-specific probability of being the father
to any age-1 offspring produced by females in the male's breeding group. The true shape and 
parameter values of $\mathrm{RV}_M$ are unknown. Here, we modeled it as a power function, and in 
the base OM its exponent was set equal to that of $\mathrm{RV}_F$ (Figure \@ref(fig:RC-plot), panel A), 
which corresponds to assuming age-specific $\mathrm{RV}_M$ equals that of the females (the current
assumption of the gag stock assessment). Mating was assumed to 
be fully random within breeding groups.

Gag life history data and age schedules (Figure S1) were 
taken from the 2022 gag stock assessment [@SEDAR2022]. To simulate total
annual survival probability at age, we followed the fisheries modelling approach 
of separating survival into age-specific, instantaneous natural
and fishing mortality components, $M$ and $F$, respectively.
Fishing mortality was further separated by fishing fleet. Traditional fisheries models
use stock-recruitment relationships to represent the density-dependent effects that
result in reduced recruitment at low spawner abundance. We instead explicitly incorporated
density-dependent age-1 survival into our simulations, such that
age-1 natural mortality had density-independent as well 
as density-dependent components. The density-dependent component was assumed to 
scale linearly with juvenile (ages 1 and 2) relative densities (relative to unfished time, $t_0$) such that 
at a relative juvenile density of 1, age-1 density-dependent survival was about 
17 percent, and at a relative juvenile density of 0.1 it was about 84 percent. 
Age-1 natural mortality was furthermore subject to random variation $\delta$ to 
help produce random recruitment variability. The total probability of survival in year
$t$ for age $a$ was calculated as:
\begin{equation}
S_{a,t} =
\begin{cases}
\begin{aligned}
&\exp\biggl[-\biggl(M_a + \max(1.75 \times \frac{\sum{N_{a\in\{1,2\},t}}}{\sum{N_{a\in\{1,2\},t_0}}}  + \delta_t, 0) + \sum_{f} v_{a,f}F_{t,f}\biggr)\biggr] & \text{if } a = 1 \\
&\exp\biggl[-\biggl(M_a + \sum_{f} v_{a,f}F_{t,f}\biggr)\biggr] & \text{if } a > 1
\end{aligned}
\end{cases}
\end{equation}
where $F_{t,f}$ is the mortality due to fleet $f$ and $v_{a,f}$ is the
vulnerability of age-$a$ fish to fleet $f$. Two fleets were 
simulated: one recreational and one commercial. Vulnerability
(i.e., the combined effect of gear selectivity 
and fish availability) for the recreational fleet was modeled as a dome-shaped 
gamma function while commercial vulnerability was specified as a logistic function 
(Figure S2, panel A), meaning the oldest individuals are only selected 
for by the commercial fleet. The simulated fishing mortality time series (Figure 
S2, panel B) followed the trend estimated by the 2022 gag
stock assessment.

Probability of reproducing in a given year was described by a logistic function, 
and the sex transition function was modeled as a cumulative normal distribution function 
(Figure S1). We simulated 40 implicitly spatially-distinct breeding groups. 
Within the spawning season, mating was restricted to individuals within the same 
breeding group. All offspring were female 
at birth and entered the population at age 1 as members of their parents’ breeding 
group. Females were allowed to switch breeding groups (see Supplement for details). Males 
did not switch breeding groups but instead remained in the breeding group of their 
sex transition year until death, consistent with the view that male gag stay on the 
spawning grounds year-round while females migrate between spawning and feeding areas.

Mortality, sex transition and whether a female reproduced in a given spawning 
season were simulated by drawing random values from uniform distributions 
between zero and one. These random draws were compared to the corresponding probabilities 
(such as the probability of a female of age $a$ reproducing), with the event (e.g., reproduction) 
occurring if the random draw was less than the probability.
Mortality source was determined for each individual 
that died in a given year. The probability of having died from natural mortality, 
commercial fishing, or recreational fishing was equal to the normalized instantaneous rates
of each mortality source for a given age class and year.

The simulation was run for 119 years, starting with an unfished population and 
ending with a highly depleted population. For each OM scenario (see below), 40
independent replicates of the population were simulated, providing 40
unique whole-population pedigrees.

### OM Scenarios

We simulated four alternative versions of 'reality' in our OM to explore the
effect of uncertainty about male contribution to reproductive success: in addition to
the base case, one scenario included sperm limitation and two scenarios
simulated (age-specific) male reproductive value functions ($\mathrm{RV}_M^\mathrm{OM}$)
that differed from those of the females ($\mathrm{RV}_M^\mathrm{EM}$). In the base OM, male 
and female reproductive value were equal, corresponding 
to what is currently assumed in the gag stock assessment, but in two alternative 
scenarios, we specified $\mathrm{RV}_M^\mathrm{OM}$ functions that increased more or less steeply
from $\mathrm{RV}_F^\mathrm{OM}$ (Figure \@ref(fig:RC-plot), panel A).

The true $\mathrm{RV}_M$ in protogynous hermaphrodites, including gag,
is unknown. It is assumed that sperm limitation exists but it has not been
measured. However, it has been shown that skipped spawning is relatively common in female gag,
and this is thought to be linked to low male abundance [@lowerre2022spawning].
In the base OM configuration, declining male abundance had no effect (meaning all eggs produced
by females will get fertilized, regardless of sex ratio). Alternately, we simulated the effect of
sperm limitation as an increased
likelihood of a female not producing any offspring in a given spawning season as the sex ratio
declines. Note that a female skipping spawning was independent between years and thus there was no systematic
skipped spawning such as alternate year spawning. The relationship between the 
current proportion of males in a spawning group relative to unfished levels, $p_M$,
and the probability of females in that spawning group skipping spawning was
modeled as an exponential function of the form $1-\exp(-{\tau p_M})$ (Figure
\@ref(fig:RC-plot), panel B), where $\tau$ controls the slope of the curve. We used a sigmoid
function to map probability of spawning to age, such that skipped spawning was more likely
in younger females. Thus, skipped spawning serves as a de facto modifier of the vector 
describing the age-specific probability of a female producing offspring in a given year.

## Sampling Model
Three types of observations were generated from the OM for use in the
integrated assessment model: CKMR kinships 
<!--
(i.e., the counts $\KHSS{c_i,c_j}{g_i,s_i}$, $\KHSD{c_i,c_j}{g_i,s_i}$,
$\KPOP{c_i,c_j}{g_i,s_i}{=}$, $\KPOP{c_i,c_j}{g_i,s_i}{\neq}$, and
$\KNot{c_i,c_j}{g_i,s_i}$), 
-->
(i.e., counts of the number of observed HSPs and POPs sharing or not sharing mtDNA
as well as the number of unrelated pairs),
a recruitment (age-1) relative abundance index, and age compositions. 
These served as data inputs for the estimation model (see below). 
Note that, unlike traditional fishery models, we did
not simulate or fit to catch data; scale for absolute abundance estimates was provided solely by
the CKMR data.

Samples for CKMR were drawn
during the final years of the simulation, from all dead individuals in the population, 
sampling age 1 for juveniles and ages 4+ for adults. In reality, adult samples would be obtained
through a combination of fishery-independent and fishery-dependent sampling, as most of the
males would need to be sampled in marine protected areas, given their scarcity in open areas [@lowerre2020testing].
The kin relationship between every pair of sampled
individuals was identified from the pedigree, and identity by descent of mtDNA between relatives
was recorded from the sex of the parent (for POPs) and of the shared parent (for HSPs).  Pairs that did not share mtDNA identical-by-descent were assigned matching
mtDNA with probability $1-H_\mathrm{mt}$. These observations were tallied according
to both individuals' birth years as well as the first-born's sampling year and sex at time of sampling.
Only pairs of samples with different birth years ("cross-cohort
comparisons") were included in the CKMR likelihood to minimize the effect of
interannual and inter-individual variation in the number of offspring [@waples2022close].
The total number of CKMR samples collected across all sampling years (see below) was chosen as various multiples of the
$10\sqrt{N}$ guideline from @bravington2016close for a single sample of a
roughly equal mix of juveniles and adults. We used the average simulated abundance of
age 2+ individuals over the CKMR sampling period as $N$ in the total sample
size calculation and the average sex ratio using all ages to partition samples between males and females.
The effect of increasing or decreasing the relative proportion of males sampled is explored in the supplement.
Thus, the sample sizes varied across simulation replicates of the OM.
The total genetic samples
were spread out over the CKMR sampling period, such that if there were fewer
sampling years, more samples per year would be collected. For each of the 40
simulated pedigrees per OM configuration, dead individuals were subsampled with a different random seed ten 
times to create ten distinct CKMR data sets per pedigree, for a total of 400 CKMR data sets per
OM configuration (the subsampling effect is explored in the Supplement).

The observed recruitment index, $I_t$, was simulated as:
\begin{equation}
I_t=\frac{N_{a=1,t}}{\max_t(N_{a=1,t})}e^{\epsilon_t}
\end{equation}
where $\epsilon$ is a normally-distributed error term with lognormal bias correction:
$\epsilon_t=\mathcal{N}(0,\sigma_I)-0.5\sigma_I^2$.
$\sigma_I$ was set to 0.01, thus producing very low variability to better
assess the uncertainty due to the CKMR and hermaphroditic aspects of the model
in this study. 
The age-1 population was indexed after density-dependent and stochastic mortality but prior to density-independent
mortality to align as closely as possible with the EM.

Age compositions were recorded as the relative proportion of each age class
in any given year, where age-1 abundance, once again, reflected pre-density-independent
mortality status. Note that we did not sample the age-composition but instead assumed it
was known perfectly. As with the unrealistically low abundance index $\epsilon$, this
was done to remove uncertainty due to fisheries data and focus solely on CKMR.

### Genetic Sampling Scenarios

To evaluate the impact of sampling on the number of observed kin pairs and 
estimator performance, we specified two
different sampling periods and three different sample sizes. In the base case, 
CKMR samples were collected over three years. Ten years
of sampling was explored as an alternative to see whether the time period over which sampling
is spread out affects the number of related pairs found and EM performance.
The base CKMR sample size, $n_\mathrm{base}$ was equal to the $10\sqrt{N}$
heuristic. The EM was additionally run with 50% and 150% of that base sample size.

## Estimation Model

We employed a standard age-structured deterministic fishery population dynamics
model and integrated the CKMR estimators derived in the Appendix to see how well the 
simulated sex-specific abundances could be estimated, given the data generated from the OM. 
Assumed known were all life history and fishery
vulnerability age schedules except sex change probability, $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$,
and age-specific male reproductive value, $\mathrm{RV}_{a,M}$. The inflection point of the two-parameter 
cumulative normal distribution $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
function was always assumed known but the standard
deviation, $\mathrm{STF}_{SD}$, which controls the function's shape was estimated in some EM runs. 
The exponent of the $\mathrm{RV}$
power function was assumed known in the EM only for females 
($\mathrm{RV}_F^\mathrm{OM} = \mathrm{RV}_F^\mathrm{EM}$, more on this below). Additional estimated parameters were mean
recruitment $\rbar$ and fleet-specific fishing mortality rates $F$ as well as annual
deviations for each. No stock-recruitment relationship was simulated or attempted
to be estimated, and unfished abundance was also not estimated since the focus of
the study was to estimate scale (i.e., absolute abundance) and sex ratio in recent years. Full equations
for the total objective function and its components are presented in Appendix B. 
The estimation model was implemented in AD Model Builder [@fournier2012ad].

### EM Scenarios

The EM for the base case included the full CKMR pseudolikelihood (for both HSPs and POPs)
and assumed that the degree of mitochondrial haplotype diversity in the population ($\Hmt$),
$\tprob_{a}$, and $\mathrm{RV}_M$ were known. 
Alternative scenarios tested models that included HSPs only in the likelihood or
contained misspecification (Figure \@ref(fig:simoverview-plot)). 
One misspecification was an incorrect $\Hmt$, which affects our ability
to correctly infer the sex of an unobserved parent or of an observed parent at a time
prior to sampling, potentially biasing CKMR kinship probabilities.
Other misspecifications were that skipped spawning was simulated 
in the OM but not accounted for in the EM and that 
$\mathrm{RV}_M^\mathrm{EM} = \mathrm{RV}_F^\mathrm{EM}$ when 
$\mathrm{RV}_M^\mathrm{OM} \neq \mathrm{RV}_F^\mathrm{OM}$. 
Finally, we allowed the EM to estimate $\mathrm{RV}_M$ to 
assess whether the model could infer this unknown quantity from the CKMR data, 
and to estimate $\tprob_{a}$ to determine if the CKMR data could allow estimating the sex ratio.

In evaluating results when HSPs only are included in the likelihood, we are
asking whether our ability to estimate abundance deteriorates when
POPs are excluded from the CKMR pseudolikelihood. In reality, POPs 
in open marine fish populations, even severely depleted ones, are going to be very rare 
and thus may not contribute much information to the model, 
so we tested the value of including POPs.

>>>>

---
title             : "Application of Close-kin mark–recapture (CKMR) to Protogynous Fishes"
subtitle          : "Figure Captions"
author            : 
  - Claudia Friess, Eric C. Anderson, Lisa E. Ailloud, David S. Portnoy, and
  - Susan K. Lowerre-Barbieri

bibliography      : ["citations_abbrev.bib", "r-references.bib"]
csl               : "elsevier-harvard.csl"

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

@@

documentclass     : article
output            : pdf_document
geometry          : margin=1in
linestretch       : 1.5
fontsize          : 12pt

@@

  - \raggedright 

@@

  - \usepackage{setspace}
  - \usepackage{caption} 
  - \setlength{\baselineskip}{1.5\baselineskip}

@@

**Figure 1.** Schematic overview of the simulation routine. The
operating model (OM) concerns the simulation of individuals in the population, with four
different scenarios listed. Observed kin pair data are obtained by sampling (sampling model) from the simulated population with different scenarios varying by the number of CKMR sampling years and 
number of samples collected. Finally, several estimation model (EM) configurations were specified. Estimated abundance and sex ratio 
were compared to simulated values by calculating percent relative error. The scenarios in each submodel highlighted by bold font are the base case.
RV = reproductive value, P(\male $\rightarrow$ \female) = sex transition probability. $\Hmts$ = simulated mitochondrial 
haplotype diversity, $\Hmte$ = mitochondrial haplotype diversity assumed in the estimation model.

\vspace{1em} 
**Figure 2.** Overview of annual steps simulated in the individual-based operating model.

\vspace{1em} 
**Figure 3.** Simulated A) reproductive value (RC) at age and B) probability of skipping spawning as a function of sex ratio. In A), the dashed line corresponds to relative female fecundity at age. The solid lines are male RC, where red is the OM scenario where $\mathrm{RV}_M = \mathrm{RV}_F$. Flatter and steeper $\mathrm{RV}_M$ are
shown in blue and green, respectively. Note that age 4 is the youngest age at which a female can turn into a male in our OM.

\vspace{1em} 
**Figure 4.** Simulated female (left) and male (right) abundance for years 75 to 119 by OM scenario (grey = base, red = skipped spawning, blue = male RC steeper than female, green = male RC flatter than females). Each of the 40 abundance time series per scenario is plotted individually, and the median is presented by bold lines.

\vspace{1em} 
**Figure 5.** Number of CKMR pairwise relationships for three years of CKMR sampling at $10\sqrt{N}$ by OM scenario. 
$\mathrm{POP}^{\neq}$ = parent-offspring pairs with different mitochondrial DNA, $\mathrm{POP}^=$ = parent-offspring pairs with matching mitochondrial DNA, 
$\mathrm{HSP}^{\neq}$ = half-sibling pairs with different mitochondrial DNA, $\mathrm{HSP}^=$ = half-sibling pairs with matching mitochondrial 
DNA. In the base OM scenario, there is no skipped spawning, male and female reproductive value (RV) are equal. The box represents the interquartile range (IQR), the line is the median, 
the whiskers extend to 1.5 IQR and outliers fall beyond that.

\vspace{1em} 
**Figure 6.** Percent relative error in estimated male and female abundance time series 
summarized across years, by number of CKMR sampling years (red = 3, blue = 10), CKMR sample size, and whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are were assumed known (fixed). Results are shown for the base OM scenario and without
any model misspecification. Y-axis truncated to improve readability.

\vspace{1em} 
**Figure 7.** Percent relative error in estimated male and female abundance time series
summarized across years, by CKMR information availability and whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are assumed known (fixed). Results are shown for the base CKMR sampling scenario. In the base EM, both half-sibling
pairs and parent-offspring-pairs were used in the likelihood and the assumed mitochondrial haplotype diversity,
$\Hmte$ (set to 1), is equal to the simulated haplotype diversity, $\Hmts$. When $\Hmte \neq \Hmts$, one was set to 1 and the other to 0.9.
Notice the different y-axis scale for panel C. Y-axis truncated to improve readability.

\vspace{1em} 
**Figure 8.** Percent relative error in estimated sex ratio 
(number of estimated males divided by total number of age 3+ year-olds) summarized across years 
under different OM and EM configurations and by whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are assumed known (fixed). In the Base case there is no skipped spawning, male and female RV are
equal, and $\Hmte = \Hmts = 1$. When $\Hmte \neq \Hmts$, one was set to 1 and the other to 0.9.
Notice the different y-axis scale for panel D. Y-axis truncated to improve readability.

\vspace{1em} 
**Figure 9.** Estimated sex transition and male reproductive value (RV) functions under different OM and EM configurations.
The dashed black lines are the simulated values and solid black lines are the median estimated values. sHmt = simulated mitochondrial haplotype diversity, eHmt = mitochondrial haplotype diversity assumed in the estimation model. In the Base case there is no skipped spawning, male and female RV are equal, and eHmt = sHmt = 1. When eHmt $\neq$ sHmt, one was set to 1 and the other to 0.9.

\vspace{1em} 
**Figure 10.** Percent relative error in estimated male and female abundance time series 
summarized across years for the OM scenarios exploring the effect of male reproductive output and by
whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are assumed known (fixed). When $P^{\mathrm{F}\rightarrow \mathrm{M}}$ was
fixed, it was fixed at the true value. When $\mathrm{RV}_{M}$ was fixed, however, it was fixed at the same
value as that of females. Consequently when $\mathrm{RV}_M$ is fixed (red and blue), it is misspecified in the EM when
$\mathrm{RV}_M^\mathrm{OM}$ differs from $\mathrm{RV}_F^\mathrm{OM}$ (panels C and D). In the base OM, there is no skipped spawning and male
and female RV are equal. Y-axis truncated to improve readability.
